---
output:
  word_document:
    reference_docx: myref.docx
---

```{r setup, include = F}

rm(list=ls())

#Load knitr
library(knitr)

#Set some options
options(na.action='na.fail')
opts_knit$set(root.dir = '..')
opts_chunk$set(echo=FALSE,
               warning=FALSE,
               message=FALSE,
               fig.path='../Figures/',
               fig.width= 8,
               fig.height= 8,
               fig.show='asis',
               dev=c('png'))

```

```{r load and clean, include = F}

#Load and clean data
source('Scripts/Telomere parentage LOAD.r')
source('Scripts/Telomere parentage FUNCTIONS.r')
source('Scripts/Telomere parentage CLEAN.r')


```


**Tables and Figures**

**Figure 1** Relationship between parental and offspring telomere length in the Seychelles warbler.

**Figure 2** Relationship between and parental age and offspring telomere length in the Seychelles warbler

**Figure 3** Parental telomere length and offspring sex ratio in the Seychelles warbler.

**Figure 4** Parental age and offspring sex ratio in the Seychelles warbler.



**Figure 1**
```{r Figure1}

Fig1A <- ggplot(subset(juv,EPP == 'Within pair'),aes(x = EmumTL, y = TL)) +
  geom_point() +
  stat_smooth(method=lm) +
  xlab('Maternal juvenile TL') +
  ylab('Offspring TL') +
  theme_lgs() +
  annotate(geom = 'text',x = 1000,y = 1.15000,label = 'A')

Fig1B <- ggplot(juv,aes(x = LmumTL, y = TL)) +
  geom_point() +
  stat_smooth(method=lm) +
  xlab('Maternal adult TL') +
  ylab('Offspring TL') +
  theme_lgs()
  annotate(geom = 'text',x = 1000,y = 15000,label = 'B')

Fig1C <- ggplot(subset(juv,EPP == 'Within pair'),aes(x = EdadTL, y = TL)) +
  geom_point() +
  stat_smooth(method=lm) +
  xlab('Paternal juvenile TL') +
  ylab('Offspring TL') +
  theme_lgs()
  annotate(geom = 'text',x = 1000,y = 15000,label = 'C')

Fig1D <- ggplot(juv,aes(x = LdadTL, y = TL)) +
  geom_point() +
  stat_smooth(method=lm) +
  xlab('Paternal adult TL') +
  ylab('Offspring TL') +
  theme_lgs()
  annotate(geom = 'text',x = 1000,y = 15000,label = 'D')

multiplot(Fig1A,Fig1B,Fig1C,Fig1D,layout=matrix(1:4,2,2))

```



**Figure 2**
```{r Figure2}

Fig2A <- ggplot(juv,aes(x = mumage, y = TL)) +
  geom_point() +
  stat_smooth(method=lm) +
  xlab('Maternal age at hatching') +
  ylab('Offspring TL') +
  theme_lgs() +
  annotate(geom = 'text',x = 1,y = 15000,label = 'A')

Fig2B <- ggplot(juv,aes(x = mumlife, y = TL)) +
  geom_point() +
  stat_smooth(method=lm) +
  xlab('Maternal lifespan') +
  ylab('Offspring TL') +
  theme_lgs() +
  annotate(geom = 'text',x = 1,y = 15000,label = 'B')

Fig2C <- ggplot(juv,aes(x = dadage, y = TL)) +
  geom_point() +
  stat_smooth(method=lm) +
  xlab('Paternal age at hatching') +
  ylab('Offspring TL') +
  theme_lgs() +
  annotate(geom = 'text',x = 1,y = 15000,label = 'C')

Fig2D <- ggplot(juv,aes(x = dadage, y = TL)) +
  geom_point() +
  stat_smooth(method=lm) +
  xlab('Paternal lifespan') +
  ylab('Offspring TL') +
  theme_lgs() +
  annotate(geom = 'text',x = 1,y = 15000,label = 'D')

multiplot(Fig2A,Fig2B,Fig2C,Fig2D,layout=matrix(1:4,2,2))

```












**Figure 3**
```{r Figure 3}

Fig3A <- ggplot(juv,aes(x = EmumTL,y = SexEstimate)) +
  geom_point(alpha=0.2) +
  stat_smooth(method = 'glm',family = 'binomial',se=T) +
  xlab('Maternal juvenile TL') +
  ylab('Proportion male offspring') +
  theme_lgs() +
  geom_abline(slope = 0,intercept = 0.5,lty=2) +  
  annotate(geom = 'text',x = 1000,y = 1.1,label = 'A')



Fig3B <- ggplot(juv,aes(x = LmumTL,y = SexEstimate)) +
  geom_point(alpha=0.2) +
  stat_smooth(method = 'glm',family = 'binomial',se=T) +
  xlab('Maternal adult TL') +
  ylab('Proportion male offspring') +
  theme_lgs() +
  geom_abline(slope = 0,intercept = 0.5,lty=2) +  
  annotate(geom = 'text',x = 1000,y = 1.1,label = 'B')


Fig3C <- ggplot(juv,aes(x = EdadTL,y = SexEstimate)) +
  geom_point(alpha=0.2) +
  stat_smooth(method = 'glm',family = 'binomial',se=T) +
  xlab('Paternal juvenile TL') +
  ylab('Proportion male offspring') +
  theme_lgs(addlegend = T) +
  geom_abline(slope = 0,intercept = 0.5,lty=2) +  
  annotate(geom = 'text',x = 1000,y = 1.1,label = 'C')



Fig3D <- ggplot(juv,aes(x = LdadTL,y = SexEstimate)) +
  geom_point(alpha=0.2) +
  stat_smooth(method = 'glm',family = 'binomial',se=T) +
  xlab('Paternal adult TL') +
  ylab('Proportion male offspring') +
  theme_lgs() +
  geom_abline(slope = 0,intercept = 0.5,lty=2) +  
  annotate(geom = 'text',x = 1000,y = 1.1,label = 'D')


multiplot(Fig3A,Fig3B,Fig3C,Fig3D,layout=matrix(1:4,2,2))


```


**Figure 4**
```{r Figure 4}

Fig4A <- ggplot(juv,aes(x = mumage,y = SexEstimate)) +
  geom_point(alpha=0.2) +
  stat_smooth(method = 'glm',family = 'binomial',se=T) +
  xlab('Maternal age at hatching') +
  ylab('Proportion male offspring') +
  theme_lgs() +
  geom_abline(slope = 0,intercept = 0.5,lty=2) +  
  annotate(geom = 'text',x = 1.1,y = 1.1,label = 'A')



Fig4B <- ggplot(juv,aes(x = mumlife,y = SexEstimate)) +
  geom_point(alpha=0.2) +
  stat_smooth(method = 'glm',family = 'binomial',se=T) +
  xlab('Maternal lifespan') +
  ylab('Proportion male offspring') +
  theme_lgs() +
  geom_abline(slope = 0,intercept = 0.5,lty=2) +  
  annotate(geom = 'text',x = 1.1,y = 1.1,label = 'B')



Fig4C <- ggplot(juv,aes(x = dadage,y = SexEstimate)) +
  geom_point(alpha=0.2) +
  stat_smooth(method = 'glm',family = 'binomial',se=T) +
  xlab('Paternal age at hatching') +
  ylab('Proportion male offspring') +
  theme_lgs() +
  geom_abline(slope = 0,intercept = 0.5,lty=2) +  
  annotate(geom = 'text',x = 1.1,y = 1.1,label = 'C')




Fig4D <- ggplot(juv,aes(x = dadlife,y = SexEstimate)) +
  geom_point(alpha=0.2) +
  stat_smooth(method = 'glm',family = 'binomial',se=T) +
  xlab('Paternal lifespan') +
  ylab('Proportion male offspring') +
  theme_lgs() +
  geom_abline(slope = 0,intercept = 0.5,lty=2) +  
  annotate(geom = 'text',x = 1.1,y = 1.1,label = 'D')



multiplot(Fig4A,Fig4B,Fig4C,Fig4D,layout=matrix(1:4,2,2))


```
