---
output:
  word_document
---
```{r setup, include = F}

rm(list=ls())

#Load knitr
library(knitr)

#Set some options
options(na.action='na.fail')
opts_knit$set(root.dir = '..')
opts_chunk$set(echo=FALSE,
               warning=FALSE,
               message=FALSE,
               fig.path='../Figures/',
               fig.width= 5,
               fig.height= 4,
               fig.show='asis',
               dev=c('png','pdf'))

```

```{r load and clean, include = F}

#Load and clean data
source('Scripts/Telomere parentage LOAD.r')
source('Scripts/Telomere parentage FUNCTIONS.r')
source('Scripts/Telomere parentage CLEAN.r')


```




**Telomere inheritence in the Seychelles warbler**


Have a look at the relationship between Parental and offspring telomere length. 

```{r Figure1,fig.height = 8}

Fig1A <- ggplot(juv,aes(x = mumTL, y = TL,col = mumAgeF)) +
          geom_point() +
          stat_smooth(method=lm) +
          xlab('Mother TL') +
          ylab('Offspring TL') +
          theme_classic()

Fig1B <- ggplot(juv,aes(x = dadTL, y = TL,col = dadAgeF)) +
          geom_point() +
          stat_smooth(method=lm) +
          xlab('Father TL') +
          ylab('Offspring TL') +
          theme_classic()

multiplot(Fig1A,Fig1B)

```


Run some models - In the first table, the same factors as before predict offspring telomere length, but so does maternal telomere length. In the second table nothing is significant. This could be collinearity, or it could be crappy sample size. Need to check.

```{r model}


lm1 <- lm(TL~mumTL+mumage+Ageclass+Sex,data = juv)
summary(lm1)$coefficients



lm2 <- lm(TL~dadTL+dadage+Ageclass+Sex,data = juv,na.action=na.exclude)
summary(lm2)$coefficients



```




Second, telomere length in both male and female parents is negatively related to the proportion of male offspring. So, mums and dads with longer telomeres have more female chicks.

```{r Figure 3,fig.width=5,fig.height=8}
juv$SexBin <- ifelse(juv$Sex == 'Females',0,1)

juv2 <- subset(juv,EPP == 'Within pair')

Fig3A <- ggplot(juv2,aes(x = mumTL,y = SexBin,col=Ageclass)) +
  geom_point(alpha=0.2) +
  stat_smooth(method = 'glm',family = 'binomial',se=F) +
  xlab('Maternal TL') +
  ylab('Proportion male offspring')


Fig3B <- ggplot(juv,aes(x = dadTL,y = SexBin,col=Ageclass)) +
  geom_point(alpha=0.2) +
  stat_smooth(method = 'glm',family = 'binomial',se=F) +
  xlab('Paternal TL') +
  ylab('Proportion male offspring')



multiplot(Fig3A,Fig3B)


```




Run some models - we get significant effects for paternal age, paternal TL, and maternal TL, but not maternal age
```{r models 2}
lm2 <- glmer(SexBin~Ageclass+dadTL*dadage+(mother|TerritoryID)+(father|TerritoryID),data=juv,family='binomial',na.action=na.exclude)
summary(lm2)


lm2 <- glmer(SexBin~mumTL*mumage +(mother|TerritoryID) + (father|TerritoryID),data=juv,family='binomial',na.action=na.exclude)
summary(lm2)


```




So is it that that males are 
