---
output:
  word_document:
    reference_docx: myref.docx
---

```{r setup, include = F}

rm(list=ls())

#Load knitr
library(knitr)

#Set some options
options(na.action='na.fail')
opts_knit$set(root.dir = '..')
opts_chunk$set(echo=FALSE,
               warning=FALSE,
               message=FALSE,
               include = FALSE,
               fig.show = 'hide')

```

```{r load and clean}

#Load and clean data
source('Scripts/Telomere parentage LOAD.r')
source('Scripts/Telomere parentage FUNCTIONS.r')
source('Scripts/Telomere parentage CLEAN.r')


```


**Tables and Figures**

**Figure 1** Juvenile sex ratio in the Seychelles warbler. **A** Sex ratio over time in juveniles. The dotted lines represent 95% confidence limits from simulated sex ratio data based on temporal variation in sample size. **B** Offspring sex ratio in relation to maternal telomere length. Individual points represent individual male (top) and female (bottom) offspring, and points are plotted with transparency so that overlap can be observed (i.e. darker points represent multiple, overlapping individuals). Lines and shaded areas represent fitted values and 95% confidence limits from a logistic regression. TL = telomere length.



**Figure 2** Maternal telomere length and offspring condition and survival in the Seychelles warbler. **A** Offspring telomere length in telation to maternal telomere length. **B** Offspring survival to adulthood in relation to telomere length. Lines and shaded areas represent estimates and 95% confidence limits from a linear regression (**A**) and logistic regression (**B**).



```{r Figure1}

Fig1A <- ggplot(ddFig1,aes(x = Year, y = sr)) +
  geom_line() +
  geom_line(aes(y = upperCI),lty = 2) +
  geom_line(aes(y = lowerCI),lty = 2) +
  theme_lgs() +
  ylim(c(0,1)) +
  ylab('Proportion male offspring') +  
  annotate(geom = 'text',x = 1990,y = 1,label = 'A')

Fig1B <- ggplot(juv2,aes(x = LmumTLKB,y = SexEstimate)) +
  geom_point(alpha=0.2) +
  stat_smooth(method = 'glm',family = 'binomial',se=T) +
  xlab('Maternal telomere length (kb)') +
  ylab('Proportion male offspring') +
  theme_lgs(addlegend = T) +
  annotate(geom = 'text',x = 1,y = 1,label = 'B')

pdf('Figures/Figure 1.pdf',
    paper = 'a4',
    width = 4,
    height =8)
multiplot(Fig1A,Fig1B)
dev.off()

```

```{r Figure2}


Fig2A <- ggplot(juv2,aes(x = LmumTLKB, y = TLKB,col = Sex, fill = Sex)) +
  geom_point() +
  stat_smooth(method=lm) +
  xlab('Maternal telomere length (kb)') +
  ylab('Offspring telomere length (kb)') +
  theme_lgs() +
  annotate(geom = 'text',x = 1,y = 12,label = 'A')

Fig2B <- ggplot(juv2,aes(x = TLKB,y = SurvivedNext,col = Sex,fill=Sex)) +
  geom_point(alpha=0.2) +
  stat_smooth(method = 'glm',family = 'binomial',se=T,fullrange=T) +
  xlab('Telomere length (kb)') +
  ylab('Prob4bility of survival') +
  theme_lgs() +
  annotate(geom = 'text',x = 1,y = 1,label = 'B')

pdf('Figures/Figure 2.pdf',
    paper = 'a4',
    width = 4,
    height = 8)
multiplot(Fig2A,Fig2B)
dev.off()
```

#Put it all into a nice table
table2 <- imptable(lmavg,myimp)
table2 <- signif(table2,2)

