---
output:
  word_document:
    reference_docx: myref.docx
bibliography: MyBib.bib
csl: Evolution.csl

---

```{r setup, include = F}

rm(list=ls())

#Load knitr
library(knitr)

#Set some options
options(na.action='na.fail')
opts_knit$set(root.dir = '..')
opts_chunk$set(echo=FALSE,
               warning=FALSE,
               message=FALSE,
               fig.path='../Figures/',
               fig.width= 5,
               fig.height= 4,
               fig.show='asis',
               dev='png')

```

```{r load and clean, include = F}

#Load and clean data
source('Scripts/Telomere parentage LOAD.r')
source('Scripts/Telomere parentage FUNCTIONS.r')
source('Scripts/Telomere parentage CLEAN.r')


```




**Adaptive sex ratio allocation is linked to maternal TL in the Seychelles warbler**

Lewis G. Spurgin^a,b^, Hannah, L. Dugdale^c,d^, Martijn Hammers, Kat Bebbington^a^, Eleanor A. Fairfield^a^, Jan Komdeur^c^, Terry Burke^d^  and David S. Richardson^a,e,^.

a. School of Biological Sciences, University of East Anglia, Norwich Research Park, NR4 7TJ, United Kingdom
b. Department of Zoology, Edward Grey Institute, University of Oxford, Oxford, UK
c. Groningen Institute for Evolutionary Life Sciences, University of Groningen, Groningen, The Netherlands
d. Department of Animal and Plant Sciences, NERC Biomolecular Analysis Facility, University of Sheffield, Sheffield, UK
e. Nature Seychelles, Roche Caiman, Mahé, Republic of Seychelles

**Correspondence:** Lewis Spurgin: lewisspurgin@gmail.com; David Richardson: david.richardson@uea.ac.uk


**Abstract**


**Introduction**

The Seychelles warbler (*Acrocephalus sechellensis*) constitutes a textbook example of adaptive sex ratio modification [@Komdeur1996; @Komdeur1997]. The Seychelles warbler is a facultative cooperative breeder - in good environmental conditions, daughters often remain on their natal territory to help rear their siblings.  Komdeur *et al.* [-@Komdeur1997] showed that 

In this study we combine telomere data with the long-term Seychelles warbler data set to test the hypothesis parental condition is related to offspring sex ratio. Specifically, we test the hypothesis that mothers and fathers with longer telomerers produce more females.



**Methods**

*Study species and sampling*

*Molecular methods*

*Statistical analyses*
We used linear mixed effects models (LMMs) to test for a relationship between parental and juvenile TL. Juvenile TL was entered as the response variable, age class as a fixed factor, and as covariates we included maternal and paternal TL, seasonal food availability, territory quality and the number of helpers present in the natal territory. As our dataset spanned many breeding seasons, and contained multiple juveniles from the same parents, we included an index of the breeding season, maternal ID and paternal ID as random factors. In order to differentiate between parental quality and parental condition, we partitioned parental TL into 'juvenile parental TL' and 'adult parental TL'. Juvenile TL is related to later-life survival, and is thus a biomarker of individual quality, whereas adult TL is expected to reflect the stresses encountered over their adult life, and is thus a marker of condition. Because juvenile and adult parental TL are correlated, we ran separate models for these two variables, keep all other variables the same.

We then separately tested whether parental age was related to offspring TL, again using LMMs. These were constructed as above, but with parental age and parental lifespan added as covriates in place of parental (adult and juvenile) TL. We chose to run these models separately from the models including parental TL due to collinearity issues between TL and age [@Barrett2012]. We compared models containing parental TL to those containing parental age using AIC values and likelihood ratio tests.

Finally, we used generalized linear mixed effects models (GLMMs) with a binomial error structure to test how parental TL is related to offspring sex ratio. These models were constructed exactly as the LMMs above, but with offspring sex as the response variable.


**Results**
```{r models,include = F}

#Juvenile TL and parental juvenile TL
juvTL_EdadTL <- lmer(TL~EdadTL+Tarsus+Helper + (1|father),data = juv,na.action=na.exclude)
juvTL_EmumTL <- lmer(TL~EmumTL+Ageclass+Tarsus+Helper + (1|mother),data = juv,na.action=na.exclude)

#Juvenile TL and parental adult TL
juvTL_LparTL <- lmer(TL~LmumTL + LdadTL + Ageclass + Tarsus + Helper + (1|mother) + (1|father),data = juv,na.action=na.exclude)


#Juvenile TL and parental age
juvTL_parage <- lmer(TL~mumage + dadage + Ageclass + Tarsus + Helper + (1|mother) + (1|father),data = juv,na.action=na.exclude)

#Juvenile TL and parental lifespan
juvTL_parlife <- lmer(TL~mumlife + dadlife + Ageclass + Tarsus + Helper + (1|mother) + (1|father),data = juv,na.action=na.exclude)


```


Seychelles warbler juvenile TL was not related to maternal juvenile TL (estimate = `r getstat(juvTL_EmumTL, 'EmumTL','est')`, CIs = `r getstat(juvTL_EmumTL, 'EmumTL','CI')`; Fig. 1A), but was positively related to maternal adult TL (estimate = `r getstat(juvTL_LparTL, 'LmumTL','est')`, CIs = `r getstat(juvTL_LparTL, 'LmumTL','CI')`; Fig. 1B). Conversely, juvenile TL was positively related to paternal juvenile TL (estimate = `r getstat(juvTL_EdadTL, 'EdadTL','est')`, CIs = `r getstat(juvTL_EdadTL, 'EdadTL','CI')`; Fig. 1C), but was not related to paternal adult TL (estimate = `r getstat(juvTL_LparTL, 'LdadTL','est')`, CIs = `r getstat(juvTL_LparTL, 'LdadTL','CI')`; Fig. 1D). Juvenile TL was not related to maternal age (estimate = `r getstat(juvTL_parage, 'mumage','est')`, CIs = `r getstat(juvTL_parage, 'mumage','CI')`; Fig. 2A), but was positvely associated with maternal lifespan (estimate = `r getstat(juvTL_parlife, 'mumlife','est')`, CIs = `r getstat(juvTL_parlife, 'mumlife','CI')`; Fig. 2B). Juvenile TL was not associated with paternal age (estimate = `r getstat(juvTL_parage, 'dadage','est')`, CIs = `r getstat(juvTL_parage, 'dadage','CI')`; Fig. 2C) or paternal lifespan (estimate = `r getstat(juvTL_parlife, 'dadlife','est')`, CIs = `r getstat(juvTL_parlife, 'dadlife','CI')`; Fig. 2D).

Sex ratio and parental TL

Sex ratio and parental age.

**Discussion**
