---
output:
  word_document:
    reference_docx: myref.docx
---

```{r setup, include = F}

rm(list=ls())

#Load knitr
library(knitr)

#Set some options
options(na.action='na.fail')
opts_knit$set(root.dir = '..')
opts_chunk$set(echo=FALSE,
               warning=FALSE,
               message=FALSE,
               include = FALSE,
               fig.show = 'hide')

```

```{r load and clean}

#Load and clean data
source('Scripts/Telomere parentage LOAD.r')
source('Scripts/Telomere parentage FUNCTIONS.r')
source('Scripts/Telomere parentage CLEAN.r')


```


**Tables and Figures**

**Figure 1** Sex ratio over time in Juvenile Seychelles warblers. The dotted and dashed lines represent equal sex ratio and 95% confidence limits from simulated sex ratio data based on temporal variation in sample size, respectively. 

**Figure 2** Parental telomere length and offspring sex ratio in the Seychelles warbler. **A** Maternal and paternal telomere length are not related to one another. **B-D** Offspring sex ratio in relation to maternal (**B**), paternal (**C**) and mean parental telomere length (**D**). Individual points represent individual male (top) and female (bottom) offspring. Lines and shaded areas represent fitted values and 95% confidence limits from a logistic regression.

**Figure 3** Parental telomere length and offspring condition and survival in the Seychelles warbler.  **A** Offspring telomere length in relation to maternal telomere length, **B** paternal telomere length, and **C** mean parental telomere length. **D** Offspring survival to adulthood in relation to telomere length. Lines and shaded areas represent estimates and 95% confidence limits from linear regressions (**A-C**) and logistic regression (**D**). Blue, dashed lines = males, red, solid lines = Females.


```{r Figure 1}
Fig1 <- ggplot(ddFig1,aes(x = Year, y = sr)) +
  geom_line() +
  geom_line(aes(y = upperCI),lty = 2) +
  geom_line(aes(y = lowerCI),lty = 2) +
  theme_lgs() +
  theme(axis.text.x = element_text(angle = 60,hjust=1)) +
  ylim(c(0,1)) +
  geom_abline(aes(intercept = 0.5,slope = 0),lty=3,col=grey(0.5)) +
  ylab('Proportion male offspring')

pdf('Figures/Figure 1.pdf',
    paper = 'a4',
    width = 3.5,
    height =3.5)
Fig1
dev.off()



```

```{r Figure2}


Fig2A <- ggplot(juv,aes(x = EdadTL,y = EmumTL)) +
  geom_point(alpha=0.2) +
  xlab('Paternal RTL (kb)') +
  ylab('Maternal RTL (kb)') +
  theme_lgs(addlegend = T)
  annotate(geom = 'text',x = 1,y = 12,label = 'A')

t1 <- tapply(juv$RTL,juv$LayYear,mean)
mymed <- median(t1)
gys <- names(which(t1>mymed))
bys <- names(which(t1<mymed))

juv$gy <- ifelse(juv$LayYear %in% gys,'gy','by')
  

temp <- FL

Fig2B <- ggplot(temp,aes(x = EmumTL,y = SexEstimate)) +
  geom_point(alpha=0.2) +
  stat_smooth(method = 'glm',family = 'binomial',se=T) +
  xlab('Maternal RTL (kb)') +
  ylab('Proportion male offspring') +
  theme_lgs(addlegend = T)
  annotate(geom = 'text',x = 1,y = 1,label = 'B')

Fig2C <- ggplot(temp,aes(x = EdadTL,y = SexEstimate)) +
  geom_point(alpha=0.2) +
  stat_smooth(method = 'glm',family = 'binomial',se=T) +
  xlab('Paternal RTL (kb)') +
  ylab('Proportion male offspring') +
  theme_lgs(addlegend = T)
  annotate(geom = 'text',x = 1,y = 1,label = 'C')
Fig2D <- ggplot(temp,aes(x = parTL,y = SexEstimate)) +
  geom_point(alpha=0.2) +
  stat_smooth(method = 'glm',family = 'binomial',se=T) +
  xlab('Mean parental RTL (kb)') +
  ylab('Proportion male offspring') +
  theme_lgs(addlegend = T)
  annotate(geom = 'text',x = 1,y = 1,label = 'D')


pdf('Figures/Figure 2.pdf',
    paper = 'a4',
    width = 7,
    height =7)
multiplot(Fig2A,Fig2B,Fig2C,Fig2D,layout=matrix(c(1,2,3,4),2,2,byrow=T))
dev.off()

```




```{r Figure3}


Fig3A <- ggplot(juv,aes(x = LmumTL, y = RTL,col = Sex, fill = Sex, lty = Sex)) +
  geom_point() +
  stat_smooth(method=lm) +
  xlab('Maternal RTL (kb)') +
  ylab('Offspring RTL (kb)') +
  theme_lgs()
  annotate(geom = 'text',x = 1.5,y = 12.5,label = 'A')
Fig3B <- ggplot(juv,aes(x = LdadTL, y = RTL,col = Sex, fill = Sex, lty = Sex)) +
  geom_point() +
  stat_smooth(method=lm) +
  xlab('Paternal RTL (kb)') +
  ylab('Offspring RTL (kb)') +
  theme_lgs()
  annotate(geom = 'text',x = 1.5,y = 12.5,label = 'B')

Fig3C <- ggplot(juv,aes(x = parTL, y = RTL,col = Sex, fill = Sex, lty = Sex)) +
  geom_point() +
  stat_smooth(method=lm) +
  xlab('Mean parental RTL (kb)') +
  ylab('Offspring RTL (kb)') +
  theme_lgs()
  annotate(geom = 'text',x = 1.5,y = 12.5,label = 'C')
Fig3D <- ggplot(FL,aes(x = RTL, y = SurvivedNext,col = Sex, fill = Sex, lty = Sex)) +
  geom_point(alpha= 0.2) +
  stat_smooth(method=glm, family = binomial) +
  xlab('Offspring RTL (kb)') +
  ylab('Survival to adulthood') +
  theme_lgs() +
  annotate(geom = 'text',x = 1,y = 1,label = 'D')


pdf('Figures/Figure 3.pdf',
    paper = 'a4',
    width = 7,
    height = 7)
multiplot(Fig3A,Fig3B,Fig3C,Fig3D,layout=matrix(1:4,2,2,byrow=T))
dev.off()
```

